Naomi model workflow

### 2022-12-01 src/example_naomi/naomi.Rmd
Working through Naomi vignette

Area hierarchy and boundaries:

@@

Population data:

@@

Survey data:

```{r}
survey_hiv_indicators <- read_csv(system.file("extdata/demo_survey_hiv_indicators.csv", package = "naomi"))
head(survey_hiv_indicators)
```
ART programmatic data:

@@

ANC programmatic data:

@@

@@

@@

@@

@@

@@

```{r}
fit <- fit_tmb(tmb_inputs)
```

```{r}
outputs <- output_package(fit, naomi_mf)
names(outputs)
```

```{r}
outputs$indicators %>%
  dplyr::filter(
    indicator == "prevalence",  # HIV prevalence
    age_group == "Y015_049"   # Age group 15-49
  ) %>%
  head()
```

@@

```{r}
system.time(outputs <- output_package(fit, naomi_mf))

@@

# Using simplified TMB template

Use the same data from above, but fit a model using our simplified Naomi TMB template, rather than the one from the package.
This may require making some edits to the simplified TMB template, such as allowing non-integer values for survey-weighted cases and sample sizes.

>>>>

Area hierarchy and boundaries, and plot to check we've got the right thing:

@@

The population data is as follows, with columns:

* `area_id`: contains the ISO, area level, and area identifier.
* `source`: where did this information come from?
* `calendar_quarter`: contains the type of year (such as CY for calendar year), the year, and the quarter
* `sex`: either female or male.
* `age_group`: in the format `Ylower_upper` where e.g. for 15-49 `lower = 15` and `upper = 49`.
* `population`: the population, which might not be an integer I guess due to survey weighting.
* `area_name`: the plain text version of the area name.
* `asfr`: age-sex fertility rate.

@@

The survey data is as follows, with columns:

* `indicator`: which survey indicator is it? Takes values `"prevalence"`, `"art_coverage"`, `"viral_suppression_plhiv"` and `"recent_infected"`.
* `survey_id`: the name of the survey, containing the country, year and type.
* `survey_mid_calendar_quarter`: which calendar quarter did the survey occur during?
* `area_id`: as before.
* `area_name`: as before.
* `res_type`: unknown.
* `sex`: as before.
* `age_group`: as before.
* `n_clusters`: the number of survey clusters in this area.
* `n_observations`: the number of individuals surveyed.
* `n_eff_kish`: the Kish effective sample size.
* `estimate`: a survey weighted estimate of the indicator.
* `std_error`: standard error of `estimate`.
* `ci_lower`: 95% lower confidence interval for `estimate`.
* `ci_upper`: 95% upper confidence interval for `estimate`.

```{r}
survey_hiv_indicators <- read_csv(system.file("extdata/demo_survey_hiv_indicators.csv", package = "naomi"))
head(survey_hiv_indicators)
```

The ART programmatic data is as follows, with columns:

* `area_id`: as before.
* `area_name`: as before.
* `sex`: as before.
* `age_group`: as before.
* `year`: the year.
* `calendar_quarter`: the calendar quarter.
* `art_current`: unknown.
* `art_new`: unknown.
* `vl_tested_12mos`: viral load tested past 12 months.
* `vl_suppressed_12mos`: viral load suppressed past 12 months.

@@

The ANC programmatic data is as follows, with columns:

@@

```{r}
pjnz <- system.file("extdata/demo_mwi2019.PJNZ", package = "naomi")
spec <- extract_pjnz_naomi(pjnz)
head(spec)
```

@@

* `scope`: A collection of `area_id`s defining the set of areas to be modelled. Usually this is simply national level, so the level 0 `area_id`.
* `level`: Area level at which to fit the model.
* `quarter_id_t1`: The first time point for the model -- approximately the midpoint of the household survey data used.
* `quarter_id_t2`: The second time point for the model -- the current time for which estimates are needed.
* `quarter_id_t3`: The third time point for the model -- the future projection for HIV estimates.

@@

## Review input data

This section is missing from the vignette!

@@

Setup the model:

@@

Prepare data inputs

@@

```{r}
tmb_inputs <- prepare_tmb_inputs(naomi_data)
fit <- fit_tmb(tmb_inputs)
```

Calculate model outputs:

```{r}
outputs <- output_package(fit, naomi_data) #' Note that in the vignette they have naomi_mf here, is this a mistake?
```

The output package consists of a data frame of indicators and metadata defining the labels for each indicator:

```{r}
names(outputs)
```

If uncertainty has not been calcualted yet, the output object retures values for the mode, but not the mean or lower and upper 95% uncertainty ranges:

```{r}
outputs$indicators %>%
  dplyr::filter(
    indicator == "prevalence",  #' HIV prevalence
    age_group == "Y015_049"   #' Age group 15-49
  ) %>%
  head()
```

The function `add_output_labels()` returns the indicators table with labels added as additional columns:

@@

Regenerate outputs with uncertainty ranges:

```{r}
system.time(outputs <- output_package(fit, naomi_data))

@@

## Plot model outputs

```{r}
indicators <- add_output_labels(outputs) %>%
  left_join(outputs$meta_area %>% select(area_level, area_id, center_x, center_y)) %>%
  sf::st_as_sf()
```

15-49 prevalence by district (which is `area_level` 4):

```{r}
indicators %>%
  filter(age_group == "Y015_049",
         indicator == "prevalence",
         area_level == 4) %>%
  ggplot(aes(fill = mode)) +
  geom_sf() +
  viridis::scale_fill_viridis(labels = scales::percent_format()) +
  th_map() +
  facet_wrap(~sex)
```

15-49 prevalence by zone (which is `area_level` 2):

```{r}
indicators %>%
  filter(age_group == "Y015_049",
         indicator == "prevalence",
         area_level == 2) %>%
  ggplot(aes(fill = mode)) +
  geom_sf() +
  viridis::scale_fill_viridis(labels = scales::percent_format()) +
  th_map() +
  facet_wrap(~sex)
```

Age-specific prevalence at a national-level:

```{r}
indicators %>%
  dplyr::filter(area_level == 0,
         sex != "both",
         age_group %in% get_five_year_age_groups(),
         calendar_quarter == "CY2018Q3",
         indicator == "prevalence") %>%
  left_join(get_age_groups()) %>%
  mutate(age_group = fct_reorder(age_group_label, age_group_sort_order)) %>%
  ggplot(aes(age_group, mean, ymin = lower, ymax = upper, fill = sex)) +
  geom_col(position = "dodge") +
  geom_linerange(position = position_dodge(0.8)) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent_format(1)) +
  facet_wrap(~area_name) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0.5))
```

ART coverage by age-sex-region:

```{r}
indicators %>%
  filter(area_level == 1,
         sex != "both",
         age_group %in% get_five_year_age_groups(),
         indicator == "art_coverage",
         calendar_quarter == "CY2018Q3") %>%
  left_join(get_age_groups()) %>%
  mutate(age_group = fct_reorder(age_group_label, age_group_sort_order)) %>%
  ggplot(aes(age_group, mean, ymin = lower, ymax = upper, fill = sex)) +
  geom_col(position = "dodge") +
  geom_linerange(position = position_dodge(0.8)) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent_format(1)) +
  facet_wrap(~area_name) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0.5))
```

Bubble plot of prevalence and PLHIV:

```{r}
indicators %>%
  filter(age_group == "Y015_064",
         area_level == 4,
         indicator %in% c("prevalence", "plhiv"),
         calendar_quarter == "CY2018Q3") %>%
  select(sex, center_x, center_y, indicator_label, mean) %>%
  spread(indicator_label, mean) %>%
  ggplot() +
  geom_sf() +
  geom_point(aes(center_x, center_y, colour = `HIV prevalence`, size = PLHIV)) +
  viridis::scale_color_viridis(labels = scales::percent_format()) +
  th_map() +
  facet_wrap(~sex)
```
